---
title: "PhenoData_TimeSeriesFit"
author: "Cazimir Kowalski"
date: "2022-10-28"
output: html_document
---

## Install and load packages
The packages bellow are necessary to visualize phenology data. Comment lines added to allow the document to knit. If you need the packages installed, uncomment the install.packages lines.
```{r}
#install.packages("readr")
#install.packages("ggplot2")
#install.packages("tidyr")
#install.packageS("dplyr")
#install.packageS("rjags")
#install.packages("rnoaa")
#install.packages("daymetr")

library(readr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(rjags)
library(rnoaa)
library(daymetr)
#devtools::install_github("EcoForecast/ecoforecastR",force=TRUE)
```

## Read in site and phenology data
Read in site data and most up to date phenology data quantified from phenocam data by the ecoforecasting initiative. We are only interested in deciduous broadleaf forests. This can be filtered in the site_data according to the pheno_cam vegetation type. The remaining sites can be used to filter the phenology data. 
```{r}

site_data <- readr::read_csv("https://raw.githubusercontent.com/eco4cast/neon4cast-targets/main/NEON_Field_Site_Metadata_20220412.csv") |> 
  dplyr::filter(phenology == 1, phenocam_vegetation == "deciduous broadleaf")
head(site_data)

NeonPheno =readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/phenology/phenology-targets.csv.gz", guess_max = 1e6) |> 
  na.omit()
head(NeonPheno)
```

## Manipulate phenology data.
1) Remove sites that are not deciduous broadleaf (sites not present in the site_data).
2) Add season column to group data based on spring green up or fall color change and leaf drop

```{r}
Pheno_decid = NeonPheno %>% #Filter sites accord to deciduous broadleaf
  filter(site_id %in% unique(site_data$field_site_id)) 

Pheno_seasons = Pheno_decid %>% #Add season variable
  mutate(season = case_when(lubridate::month(datetime) <= 6 ~ "spring",
                            lubridate::month(datetime) >= 7 ~ "fall")) 
```

## Visualize data
Use ggplot to visualize rcc and gcc data during spring green up and fall color change and leaf drop. Can plot across years or choose a single year

```{r, fig.width=10,fig.height=11}
Pheno_seasons %>% 
  filter(lubridate::year(datetime)== 2021) %>% #choosing 2021
  filter(season == "spring") %>% #spring
  filter(variable == "gcc_90") %>% #only interested in greeness 
  ggplot(aes(x=datetime,y=observation)) +
  geom_point(color="#228b22",size=0.75) + 
  theme_classic() + 
  labs(title="Spring phenology", x = "Date", y= "gcc_90") +
  facet_wrap(~site_id,ncol=5) + 
  theme(strip.background=element_blank(),panel.spacing=unit(1,"lines"))

Pheno_seasons %>% 
  filter(lubridate::year(datetime)== 2021) %>% #2021 YEAR
  filter(season == "fall") %>% # Fall
  ggplot(aes(x=datetime,y=observation,shape=variable)) + #Plot greeness and redness
  geom_point(aes(color=variable),size=0.75) + 
  scale_color_manual(values = c("#daa520","#228b22")) + 
  theme_classic() + 
  facet_wrap(~site_id,ncol=5) +
  labs(title="Fall phenology", x = "Date", y= "Reflectance value",color="Reflectance index",shape="Reflectance index") + 
  theme(strip.background=element_blank(),panel.spacing = unit(1,"lines"))
```

## Fit a historical time series using a Bayesian state-space model
Code derived from exercise six of the Ecological Forecasting 
```{r}
#label data
levels(as.factor(Pheno_decid$site_id))
GRSM_pheno = NeonPheno %>% filter(site_id=="GRSM")
time = (subset(GRSM_pheno, GRSM_pheno$variable == 'gcc_90'))$datetime
y = (subset(GRSM_pheno, GRSM_pheno$variable == 'gcc_90'))$observation


#define model
RandomWalk = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(x[t-1],tau_add) #Dependent on observation at previous time point
  }
  
  #### Initial condition
  x[1] ~ dnorm(x_ic,tau_ic) 
  
  #### Priors
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"

data <- list(y=y,n=length(y),           ## data
             x_ic=log(1000),tau_ic=100, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
)

#set parameters for mcmc
nchain = 3
init <- list()
for(i in 1:nchain){
  y.samp = sample(y,length(y),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(log(y.samp))),  ## initial guess on process precision
                    tau_obs=5/var(log(y.samp)))        ## initial guess on obs precision
}

#actually define model as a variable
j.model   <- jags.model (file = textConnection(RandomWalk),
                         data = data,
                         inits = init,
                         n.chains = 3)


## burn-in

jags.burn   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs"),
                            n.iter = 1000)
plot(jags.burn)


jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
plot(jags.out[,c(1,2)]) #Plot after burn in


#plotting the results
time.rng = c(1,length(time))       ## adjust to zoom in and out
out.grsm <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols.grsm <- grep("^x",colnames(out.grsm)) ## grab all columns that start with the letter x
ci.grsm <- apply(out.grsm[,x.cols.grsm],2,quantile,c(0.025,0.5,0.975)) ## model was fit on log scale


plot(time,ci.grsm[2,],type='n',ylim=c(0.2,0.6),ylab="GCC_90",xlim=time[time.rng],main="GRSM",xlab="Date")

## adjust x-axis label to be monthly if zoomed
if(diff(time.rng) < 100){ 
  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), format = "%Y-%m")
}
ecoforecastR::ciEnvelope(time,ci.grsm[1,],ci.grsm[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(time,y,pch="+",cex=0.5)
```

```{r}

pheno.decid.gcc90 = Pheno_decid %>% filter(variable=="gcc_90")
pheno.wide = pheno.decid.gcc90 %>% pivot_wider(names_from="site_id",values_from="observation")
pheno.2021 = pheno.wide %>% filter(datetime>'2020-12-31')
time = pheno.2021$datetime
pheno.2021 = pheno.2021[,-c(1,2)]

RandomWalk_multiplesites = "model{
  
  #loop over all sites
  for (s in 1:ns){
    
    #### Data Model
    for(i in 1:n){
      y[i,s] ~ dnorm(x[i,s],tau_obs)
    }
  
    #### Process Model
    for(i in 2:n){
      x[i,s]~dnorm(x[i-1,s],tau_add)
    }
  
  
  ## initial condition
  x[1,s] ~ dnorm(x_ic,tau_ic)
  }
  
  #### Priors
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}"



data <- list(y=as.matrix(pheno.2021),n=nrow(pheno.2021),
             ns = ncol(pheno.2021),
             x_ic=log(1000),tau_ic=100, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
)
#set parameters for mcmc
nchain = 3
init <- list()
for(i in 1:nchain){
  y.samp = sample(y,length(y),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(log(y.samp))),  ## initial guess on process precision
                    tau_obs=5/var(log(y.samp)))        ## initial guess on obs precision
}

#actually define model as a variable
s.model   <- jags.model (file = textConnection(RandomWalk_multiplesites),
                         data = data,
                         inits = init,
                         n.chains = 3)

jags.s.burn   <- coda.samples (model = s.model,
                            variable.names = c("tau_add","tau_obs"),
                            n.iter = 1000)
plot(jags.s.burn)
jags.s.out   <- coda.samples (model = s.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
plot(jags.s.out[,c(1,2)])

```
```{r}
#### Helper function to parse JAGS variable names that include matrix syntax (e.g. "x[40,13]")
##' @param w mcmc object containing matrix outputs
##' @param pre prefix (variable name) for the matrix variable to be extracted
##' @param numeric boolean, whether to coerce class to numeric
parse.MatrixNames <- function(w, pre = "x", numeric = FALSE) {
  w <- sub(pre, "", w)
  w <- sub("[", "", w, fixed = TRUE)
  w <- sub("]", "", w, fixed = TRUE)
  w <- matrix(unlist(strsplit(w, ",")), nrow = length(w), byrow = TRUE)
  if (numeric) {
    class(w) <- "numeric"
  }
  colnames(w) <- c("row", "col")
  return(as.data.frame(w))
} # parse.MatrixNames

out.s <- as.matrix(jags.s.out)
x.cols.s = which(substr(colnames(out.s),1,1)=="x")   ## which columns are the state variable x
ci.s <- apply(out.s[,x.cols.s],2,quantile,c(0.025,0.5,0.975)) #Credible interval
ci.names.s = parse.MatrixNames(colnames(ci.s),numeric=TRUE)
ci.t = t(ci.s) #Transpose credible intervals to parse visually if needed. 

#layout(matrix(1:15,3,5)) ## arrange plots in a 5 x 3 matrix
## plot sample of GCC_90 timeseries
for(i in 1:data$ns){
  sel = which(ci.names.s$col == i)
  plot(time,ci.t[sel,2],type='n',ylim=c(-0.25,1),ylab="gcc_90",main=colnames(data$y)[i],xlab="Date")
  ecoforecastR::ciEnvelope(x=time,ylo=ci.t[sel,1],yhi=ci.t[sel,3],col="lightBlue") #Something odd is happening here and switching the color of the credible intervals: light blue, and then white just before 2022 starts
  points(time,data$y[,i],pch="+",cex=1.5)
}
  

```







