---
title: "PhenoData_TimeSeriesFit"
author: "Cazimir Kowalski"
date: "2022-10-28"
output: html_document
---

## Install and load packages
The packages bellow are necessary to visualize phenology data. Comment lines added to allow the document to knit. If you need the packages installed, uncomment the install.packages lines.
```{r}
#install.packages("readr")
#install.packages("ggplot2")
#install.packages("tidyr")
#install.packageS("dplyr")
#install.packageS("rjags")
#install.packageS("rnoaa")
#install.packageS("daymetr")

library(readr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(rjags)
#library(rnoaa)
library(daymetr)
#devtools::install_github("EcoForecast/ecoforecastR",force=TRUE)
```

## Read in site and phenology data
Read in site data and most up to date phenology data quantified from phenocam data by the ecoforecasting initiative. We are only interested in deciduous broadleaf forests. This can be filtered in the site_data according to the pheno_cam vegetation type. The remaining sites can be used to filter the phenology data. 
```{r}

site_data <- readr::read_csv("https://raw.githubusercontent.com/eco4cast/neon4cast-targets/main/NEON_Field_Site_Metadata_20220412.csv") |> 
  dplyr::filter(phenology == 1, phenocam_vegetation == "deciduous broadleaf")
head(site_data)

NeonPheno =readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/phenology/phenology-targets.csv.gz", guess_max = 1e6) |> 
  na.omit()
head(NeonPheno)
```

## Manipulate phenology data.
1) Remove sites that are not deciduous broadleaf (sites not present in the site_data).
2) Add season column to group data based on spring green up or fall color change and leaf drop

```{r}
Pheno_decid = NeonPheno %>% #Filter sites accord to deciduous broadleaf
  filter(site_id %in% unique(site_data$field_site_id)) 

Pheno_seasons = Pheno_decid %>% #Add season variable
  mutate(season = case_when(lubridate::month(datetime) <= 6 ~ "spring",
                            lubridate::month(datetime) >= 7 ~ "fall")) 
```

## Visualize data
Use ggplot to visualize rcc and gcc data during spring green up and fall color change and leaf drop. Can plot across years or choose a single year

```{r, fig.width=10,fig.height=11}
Pheno_seasons %>% 
  filter(lubridate::year(datetime)== 2021) %>% #choosing 2021
  filter(season == "spring") %>% #spring
  filter(variable == "gcc_90") %>% #only interested in greeness 
  ggplot(aes(x=datetime,y=observation)) +
  geom_point(color="#228b22",size=0.75) + 
  theme_classic() + 
  labs(title="Spring phenology", x = "Date", y= "gcc_90") +
  facet_wrap(~site_id,ncol=5) + 
  theme(strip.background=element_blank(),panel.spacing=unit(1,"lines"))

Pheno_seasons %>% 
  filter(lubridate::year(datetime)== 2021) %>% #2021 YEAR
  filter(season == "fall") %>% # Fall
  ggplot(aes(x=datetime,y=observation,shape=variable)) + #Plot greeness and redness
  geom_point(aes(color=variable),size=0.75) + 
  scale_color_manual(values = c("#daa520","#228b22")) + 
  theme_classic() + 
  facet_wrap(~site_id,ncol=5) +
  labs(title="Fall phenology", x = "Date", y= "Reflectance value",color="Reflectance index",shape="Reflectance index") + 
  theme(strip.background=element_blank(),panel.spacing = unit(1,"lines"))
```

## Fit a historical time series using a Bayesian state-space model
Code derived from exercise six of the Ecological Forecasting 
```{r}
#label data
time = (subset(NeonPheno, NeonPheno$variable == 'gcc_90'))$datetime
y = (subset(NeonPheno, NeonPheno$variable == 'gcc_90'))$observation

#define model
RandomWalk = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(x[t-1],tau_add)
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"

data <- list(y=y,n=length(y),      ## data
             x_ic=log(1000),tau_ic=100, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
)

#set parameters for mcmc
nchain = 3
init <- list()
for(i in 1:nchain){
  y.samp = sample(y,length(y),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(log(y.samp))),  ## initial guess on process precision
                    tau_obs=5/var(log(y.samp)))        ## initial guess on obs precision
}

#actually define model as a variable
j.model   <- jags.model (file = textConnection(RandomWalk),
                         data = data,
                         inits = init,
                         n.chains = 3)


## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs"),
                            n.iter = 1000)
plot(jags.out)

#only running the mcmc w 1000 iterations for now as it crashes if you try to do 10000

#plotting the results
time.rng = c(1,length(time))       ## adjust to zoom in and out
out <- as.matrix(jags.out)         ## convert from coda to matrix  
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci = matrix(NA, 2, 3)
ci[1,] <- c(quantile(out[,1], c(0.025,0.5,0.975)))
ci[2,] = c(quantile(out[,2], c(0.025,0.5,0.975)))

plot(time,ci[2,],type='n',ylim=range(y,na.rm=TRUE),ylab="Flu Index",log='y',xlim=time[time.rng])
## adjust x-axis label to be monthly if zoomed
if(diff(time.rng) < 100){ 
  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), format = "%Y-%m")
}
ecoforecastR::ciEnvelope(time,ci[1,],ci[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(time,y,pch="+",cex=0.5)
```







